---
title: "Week 2 Exercises"
author: "Annelise Lobo"
date: "July 6, 2023"
output: pdf_document
---

Please complete all exercises below. You may use stringr, lubridate, or the forcats library. 

Place this at the top of your script:
library(stringr)
library(lubridate)
library(forcats)

# Exercise 1
Read the sales_pipe.txt file into an R data frame as sales. 

```{r}
# Your code here
library(stringr)
library(lubridate)
library(forcats)

sales_pipe = read.delim("/Users/axtonpulliam/Documents/DSE5002/Week_2/Data/sales_pipe.txt",   stringsAsFactors=FALSE, sep ="|")
```

# Exercise 2
You can extract a vector of columns names from a data frame using the colnames() function. Notice the first column has some odd characters. Change the column name for the FIRST column in the sales date frame to Row.ID. 

**Note: You will need to assign the first element of colnames to a single character.**

```{r}
# Your code here
colnames(sales_pipe) [1] = "Row.ID"
```

# Exercise 3
Convert both Order.ID and Order.Date to date vectors within the sales data frame. What is the number of days between the most recent order and the oldest order?
How many years is that?
How many weeks?

**Note: Use lubridate**

```{r}
# Your code here

#for Order.Date to be converted to date vector
str(sales_pipe$Order.Date) #find structure of item
inherits(sales_pipe$Order.Date,  c("Date")) #find out if it is a date
sales_pipe$Order.Date = as.Date(sales_pipe$Order.Date, format = '%m/%d/%Y') #convert  column to date format month, day, year

#to convert ship.date to date vectors
str(sales_pipe$Ship.Date)
inherits(sales_pipe$Ship.Date, c("Date"))
sales_pipe$Ship.Date = as.Date(sales_pipe$Ship.Date, format = '%B %d %Y') 

#days between most recent order
oldest = min(sales_pipe$Order.Date) #find oldest date
recent = max(sales_pipe$Order.Date) #find newest date

day_diff = difftime(recent, oldest, units = "days") #find number of days b/w oldest and newest order
day_diff
time_length(day_diff, "year") #print the number of years difference
difftime(recent, oldest, units = "weeks") #find number of weeks b/w oldest and newest order


```


# Exercise 4
What is the average number of days it takes to ship an order?

```{r}
# Your code here
mean(difftime(sales_pipe$Ship.Date, sales_pipe$Order.Date, units = "days")) #find the average of the difference between ship and order date
```


# Exercise 5
How many customers have the first name Bill?
You will need to split the customer name into first and last name segments and then use a regular expression to match the first name bill. Use the length() function to determine the number of customers with the first name Bill in the sales data. 

```{r}
# Your code here

#Create two new columns for the first and last names based off the customer.name column
temp_char  = stringr::str_split_fixed(string = sales_pipe$Customer.Name, pattern = " ", n = 2) #split the customer.name column into two segments, seperated by a space
sales_pipe$Customer.First.Name = paste(temp_char[,1], sep=" ") #paste the first column of customer.name into a new column. These will be the first names
sales_pipe$Customer.Last.Name = paste(temp_char[,2], sep=" ") #paste the ssecond column of customer.name into a new colum. These will be the last names

#Find customers with the last name Bill
length(which(sales_pipe$Customer.First.Name=='Bill')) #finds all the customers with first name Bill, then counts how many there are
```

# Exercise 6
How many mentions of the word 'table' are there in the Product.Name column? 
**Note you can do this in one line of code**

```{r}
# Your code here
length(which(sales_pipe$Product.Name=='table'))
```

# Exercise 7
Create a table of counts for each state in the sales data. The counts table should be ordered alphabetically from A to Z. 

```{r}
# Your code here
table(sales_pipe$State)
```

# Exercise 8
Create an alphabetically ordered barplot for each sales Category in the State of Texas. 

```{r}
# Your code here
barplot(table(sales_pipe$Category), subset = sales_pipe$State == "Texas")

```

# Exercise 9
Find the average profit by region.
**Note: You will need to use the aggregate() function to do this. To understand how the function works type ?aggregate in the console.**

```{r}
# Your code here
aggregate(sales_pipe$Profit, list(sales_pipe$Region), mean)
```

# Exercise 10
Find the average profit by order year. 
**Note: You will need to use the aggregate() function to do this. To understand how the function works type ?aggregate in the console.**

```{r}
# Your code here

#split the order date and create a new column with just the order year
temp_char  = stringr::str_split_fixed(string = sales_pipe$Order.Date, pattern = "-", n =3) #split the Order.Date column into three segments, seperated by a -
sales_pipe$Order.Year = paste(temp_char[,1], sep="-") #paste the first segment of the split column into a new column named Order.Year

#Find the mean of the profits seperated by order year
aggregate(sales_pipe$Profit, list(sales_pipe$Order.Year), mean)

```


